```{r setup}
#| message: false
#| warning: false
library(DiagrammeR)
```

# Conceptual framework {.unnumbered}

## Context and motivation

-   Selectively logged forests are becoming the majority of tropical forests $\rightarrow$ Importance of understanding their recovery process;

-   Disturbances such as logging have different effects on the demographic rates of species and can therefore change the species composition and biodiversity in the long term;

-   Demographic processes can therefore provide insights into forest resilience.

## Research questions

i.  What is the contribution of demographic processes (growth, recruitment, mortality) to post-logging changes in biomass, biodiversity and composition?

ii. What is the link between biomass and biodiversity recovery processes?

iii. How do abiotic factors shape demographic processes of recovery?

## General diagram 

```{r}
grViz("
digraph {

  rankdir=LR

  graph [overlap = true, fontsize = 10]

  subgraph cluster_ba{
    label='Basal area'; fontsize = 14;
    node [shape = box, fillcolor='coral', style=filled, fontsize=12]
    ba_stock [label='Stock']
    ba_in [label='Gain']
    ba_out [label='Loss']
  }

  node [shape = box, fillcolor='coral', style=filled]
  {rank=same
  cwm [label='Functional\ncomposition']
  div [label='Functional\ndiversity']
  }
  node [fillcolor='powderblue']
  env [label='Environmental\nvariables']
  intensity [label='Logging\nintensity']

  node [shape = ellipse, fillcolor='whitesmoke']
  {rank=same
  closure[label='Canopy\nclosure']
  vmax[label='Asymptotic\nparameters*']
  }

  intensity -> closure
  closure -> {ba_in cwm}
  vmax -> {ba_out ba_in cwm div}
  env -> {closure vmax}
  cwm -> {ba_in ba_out div}
  ba_stock -> closure

  edge [style = bold]
  div -> ba_in

  edge [style = dashed]
  ba_in -> ba_stock
  ba_out -> ba_stock

  node [shape = plaintext, fillcolor='white']
  legend[label='*Asymptotic stocks: carrying capacity
  or fluxes: equilibrium turnover rate']
}
")
```
