```{r setup}
#| message: false
#| warning: false
library(tidyverse)
library(knitr)
```

## Choice of variables

Below we show the productivity in four sites that have been remeasured for multiple decades.

```{r}
data <- read.csv("data/derived_data/aggregated_data_v8.csv") 
plot_info <- read.csv("data/raw_data/bioforest-plot-information.csv")
```

```{r}
#| message: false
#| warning: false
data |>
  left_join(plot_info, by = join_by(Site == site, Plot == plot)) |>
  subset(variable %in% c("agb_growth", "agb_recr")) |>
  subset(Site %in% c("Mbaiki", "Paracou", "Ulu Muda", "Lesong")) |>
  pivot_wider(names_from = "variable") |>
  ggplot(aes(Year, agb_recr + agb_growth, ,
    group = Plot,
    col = Treatment
  )) +
  geom_point() +
  geom_line() +
  labs(y = "AGB productivity [Mg/ha/yr]", col = NULL) +
  facet_wrap(~Site, scales = "free") +
  theme_minimal()
```

```{r}
data |>
  left_join(plot_info, by = join_by(Site == site, Plot == plot)) |>
  subset(variable == "nstem_recr") |>
  subset(Site %in% c("Mbaiki", "Paracou", "Ulu Muda", "Lesong")) |>
  ggplot(aes(Year, value, group = Plot, col = Treatment)) +
  geom_point() +
  geom_line() +
  labs(y = "Stem recruitment [/ha/yr]", col = NULL) +
  facet_wrap(~Site, scales = "free") +
  theme_minimal()
```

```{r}
#| message: false
#| warning: false
data |>
  left_join(plot_info, by = join_by(Site == site, Plot == plot)) |>
  subset(variable == "agb_mort") |>
  filter(Treatment == "Control" | Year >= Year_of_harvest+3) |>
  subset(Site %in% c("Mbaiki", "Paracou", "Ulu Muda", "Lesong")) |>
  ggplot(aes(Year, value, group = Plot, col = Treatment)) +
  geom_point() +
  geom_line() +
  labs(y = "AGB mortality [Mg/ha/yr]", col = NULL) +
  facet_wrap(~Site, scales = "free") +
  theme_minimal()
```

```{r}
data |>
  left_join(plot_info, by = join_by(Site == site, Plot == plot)) |>
  subset(variable == "nstem_mort") |>
  filter(Treatment == "Control" | Year >= Year_of_harvest+3) |>
  subset(Site %in% c("Mbaiki", "Paracou", "Ulu Muda", "Lesong")) |>
  ggplot(aes(Year, value, group = Plot, col = Treatment)) +
  geom_point() +
  geom_line() +
  labs(y = "Stem mortality [/ha/yr]", col = NULL) +
  facet_wrap(~Site, scales = "free") +
  theme_minimal()
```

## Site selection

In this analysis, we only kept sites that meet the following criteria:

-   at least 4 post-logging measurements

-   

    > 10 years of post-logging measurements

-   presence of control plots

-   productivity values \< 10% (too high)?

```{r}
#| eval: false
criteria_site <- data |>
  inner_join(plot_info, by = join_by(Site == site, Plot == plot)) |> 
  group_by(Site) |> 
  summarise(controls = length(unique(Plot[Treatment=="Control"])), 
            logged = length(unique(Plot[Treatment=="Logging"])))
data |>
  inner_join(plot_info, by = join_by(Site == site, Plot == plot)) |> 
  filter(Treatment == "Logging" & !is.na(Year)) |> 
  group_by(Site, Plot) |> 
  summarise(ncensus = length(unique(Year[Year > Year_of_harvest])), 
            tcensus = max(Year, na.rm = TRUE) - unique(Year_of_harvest)) |>
  group_by(Site) |> 
  summarise(ncensus = median(ncensus), tcensus = median(tcensus)) |> 
  left_join(criteria_site) |> 
  filter(ncensus>3,tcensus>9,controls>0,logged>0) |> 
  write.csv("data/derived_data/sites_to_keep.csv", row.names = FALSE)
```

```{r}
read.csv("data/derived_data/sites_to_keep.csv") |> kable()
```

