```{r setup}
#| message: false
#| warning: false
library(tidyverse)
library(knitr)
```

# Flux data {.unnumbered}

## Choice of variables

Below we show the productivity in four sites that have been remeasured for multiple decades.

```{r}
#| message: false
#| warning: false
plot_info <- read.csv("data/raw_data/bioforest-plot-information.csv") |>
  left_join(read.csv("data/raw_data/plot_correspondance.csv")) |>
  select(site, plot = new_plot, Year_of_harvest, Treatment) |>
  unique()

data <- read.csv("data/derived_data/aggregated_data_v10.csv") |>
  left_join(plot_info, by = join_by(Site == site, Plot == plot)) |>
  filter(grepl("^agb", variable) & !grepl("]$", variable)) |>
  pivot_wider(names_from = "variable", values_from = "value") |>
  mutate(prod = agb_growth + agb_recr, mort = agb_mort, stocks = agb) |>
  select(!contains("agb")) |>
  group_by(Site, plot_new = Plot) |>
  arrange(Year) |>
  mutate(tdiff = c(0, diff(Year))) |>
  ungroup() |>
  pivot_longer(cols = c("stocks", "prod", "mort"), names_to = "variable") |>
  mutate(Year_of_harvest = ifelse(Site == "Mbaiki" & Treatment == "Logging",
    1985, Year_of_harvest
  )) |>
  mutate(Year_of_harvest = ifelse(Site == "Paracou" & Treatment == "Logging",
    1987, Year_of_harvest
  )) |>
  mutate(trec = Year - Year_of_harvest) |>
  filter(!(Site == "Tapajos Km67" &
             grepl("loggedtwice", Plot) & Year > 2014)) |>
  filter(!(Site == "Tapajos Km67" & grepl("later", Plot)))
```

```{r}
#| message: false
#| warning: false
data |>
  filter(variable == "prod") |>
  filter(Site %in% c("Paracou", "Mbaiki", "SUAS", "Lesong")) |>
  filter(Year > Year_of_harvest | Treatment == "Control") |>
  ggplot(aes(Year, value, group = plot_new, col = Treatment)) +
  geom_point() +
  geom_line() +
  labs(y = "AGB productivity [Mg/ha/yr]", col = NULL) +
  facet_wrap(~Site, scales = "free") +
  theme_minimal()
```

```{r}
#| message: false
#| warning: false
data |>
  filter(variable == "mort") |>
  filter(Site %in% c("Paracou", "Mbaiki", "SUAS", "Lesong")) |>
  filter(Year > Year_of_harvest + 3 | Treatment == "Control") |>
  ggplot(aes(Year, value, group = plot_new, col = Treatment)) +
  geom_point() +
  geom_line() +
  labs(y = "AGB mortality [Mg/ha/yr]", col = NULL) +
  facet_wrap(~Site, scales = "free") +
  theme_minimal()
```

## Site selection

In this analysis, we only kept sites that meet the following criteria:

-   at least 4 post-logging measurements

-   at least 10 years of post-logging measurements

-   presence of control plots

-   productivity values \< 10% (too high)?

```{r}
#| message: false
#| warning: false
#| eval: false
criteria_site <- data |>
  group_by(Site) |>
  summarise(
    controls = length(unique(plot_new[Treatment == "Control"])),
    logged = length(unique(plot_new[Treatment == "Logging"]))
  )
data |>
  filter(Treatment == "Logging" & !is.na(Year)) |>
  group_by(Site, plot_new) |>
  summarise(
    ncensus = length(unique(Year[Year > Year_of_harvest])),
    tcensus = max(Year, na.rm = TRUE) - unique(Year_of_harvest)
  ) |>
  group_by(Site) |>
  summarise(ncensus = median(ncensus), tcensus = median(tcensus)) |>
  left_join(criteria_site) |>
  filter(ncensus > 3, tcensus > 9, controls > 0, logged > 0) |>
  write.csv("data/derived_data/sites_to_keep.csv", row.names = FALSE)
```

```{r}
#| message: false
#| warning: false
read.csv("data/derived_data/sites_to_keep.csv") |>
  arrange(-ncensus) |>
  kable()
```
