```{r setup}
#| message: false
#| warning: false
library(tidyverse)
```

## Mortality {.unnumbered}

Due to its high stochasticity, it is challenging to identify any general trend in mortality. To avoid overcomplicating the model, two simple assumptions can be considered:

i.  the absolute mortality remains constant over time;

ii. the mortality rate (mortality/stocks) remains constant over time.

The second hypothesis has the advantage of being more robust to limit conditions (e.g. when the stocks tend towards zero, the absolute mortality cannot be higher than the stocks).

```{r}
data <- read.csv("data/derived_data/aggregated_data.csv")

control_string <- "unlogged|to be logged|old-growth|control|natural"
silv_string <- "treatment|devital|thinning"

plot_info <-
  read.csv("data/raw_data/bioforest-plot-information.csv") |>
  mutate(site = ifelse(site == "BAFOG", "Bafog", site)) |>
  mutate(site = ifelse(site == "Montagne_tortue", "Montagne_Tortue", site)) |>
  # harmonize treatment information as logging and silvicultural treatments
  mutate(
    Treatment = ifelse(Treatment != "", Treatment, NA),
    logging = !grepl(control_string, tolower(Treatment)),
    silv_treat = grepl(silv_string, tolower(Treatment)),
    year_log = ifelse(site == "Paracou", 1986,
      Year_of_harvest |> substr(1, 4) |> as.numeric()
    )
  ) |>
  select(site, plot, logging, silv_treat, year_log)
```

```{r}
#| message: false
#| warning: false
data |>
  merge(plot_info, by.x = c("Site", "Plot"), by.y = c("site", "plot")) |>
  subset(variable %in% c("ba_mort", "ba")) |>
  subset(Site %in% c("Mbaiki", "Paracou", "Ulu Muda", "Misiones")) |>
  subset((Year > year_log + 5 & !is.na(year_log)) | !logging) |>
  pivot_wider(names_from = "variable") |>
  arrange(Year) |>
  ggplot(aes(ba, ba_mort)) +
  geom_point(aes(col = logging, group = Plot)) +
  geom_path(aes(col = logging, group = Plot)) +
  geom_smooth() +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap(~Site, scales = "free") +
  theme_minimal()
```

```{r}
#| message: false
#| warning: false
data |>
  merge(plot_info, by.x = c("Site", "Plot"), by.y = c("site", "plot")) |>
  subset(variable %in% c("ba_mort", "ba")) |>
  subset(Site %in% c("Mbaiki", "Paracou", "Ulu Muda", "Misiones")) |>
  subset((Year > year_log + 5 & !is.na(year_log)) | !logging) |>
  pivot_wider(names_from = "variable") |>
  arrange(Year) |>
  ggplot(aes(Year, ba_mort / ba)) +
  geom_point(aes(col = logging, group = Plot)) +
  geom_path(aes(col = logging, group = Plot)) +
  labs(y = "Relative mortality = BAmort/BA") +
  # geom_smooth() +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap(~Site, scales = "free") +
  theme_minimal()
```

There doesn't seem to be a strong correlation between BA mortality and BA in our data, but as a first approximation considering that mortality is proportional to basal area might work.
