```{r setup}
#| message: false
#| warning: false
#| include: false

library(tidyverse)
library(cmdstanr)
library(bayesplot)
library(patchwork)
library(flextable)
library(posterior)
list.files("r", full.names = TRUE, pattern = ".R") |>
  lapply(source)
```

# Model calibration {.unnumbered}

## Data

```{r}
#| message: false
#| warning: false
plot_info <- read.csv("data/raw_data/bioforest-plot-information.csv") |>
  left_join(read.csv("data/raw_data/plot_correspondance.csv")) |>
  select(site, plot = new_plot, Year_of_harvest, Treatment) |>
  unique()

data <- read.csv("data/derived_data/aggregated_data_v10.csv") |>
  left_join(plot_info, by = join_by(Site == site, Plot == plot)) |>
  filter(grepl("^agb", variable) & !grepl("]$", variable)) |>
  pivot_wider(names_from = "variable", values_from = "value") |>
  mutate(prod = agb_growth + agb_recr, mort = agb_mort, stocks = agb) |>
  select(!contains("agb")) |>
  group_by(Site, plot_new = Plot) |>
  arrange(Year) |>
  mutate(tdiff = c(0, diff(Year))) |>
  ungroup() |>
  pivot_longer(cols = c("stocks", "prod", "mort"), names_to = "variable") |>
  mutate(Year_of_harvest = ifelse(Site == "Mbaiki" & Treatment == "Logging",
    1985, Year_of_harvest
  )) |>
  mutate(Year_of_harvest = ifelse(Site == "Paracou" & Treatment == "Logging",
    1987, Year_of_harvest
  )) |>
  mutate(trec = Year - Year_of_harvest) |>
  filter(!(Site == "Tapajos Km67" &
             grepl("loggedtwice", Plot) & Year > 2014)) |>
  filter(!(Site == "Tapajos Km67" & grepl("later", Plot)))
```

For now, we only calibrate the model with sites that meet all the criteria listed [here](https://bioforest-project.github.io/demography/21_flux_data.html), and that have at least 10 censuses.

```{r}
#| message: false
#| warning: false
sites <- read.csv("data/derived_data/sites_to_keep.csv") |>
  filter(ncensus >= 10)
data <- data |>
  filter(Site %in% sites$Site)
```

## Model

We used the following parameter bounds when calibrating the model.

```{r}
#| message: false
#| warning: false
bounds <- data.frame(
  "theta_min" = 50,
  "theta_max" = 1000,
  "lambda_min" = 1e-2,
  "lambda_max" = 5,
  "alpha_min" = 0,
  "alpha_max" = 200,
  "delta_min" = 0,
  "delta_max" = 20,
  "tau_min" = 2,
  "tau_max" = 30,
  "k_min" = 0.1,
  "k_max" = 5,
  "omega_min" = 1e-3,
  "omega_max" = 0.5
)
bounds |>
  pivot_longer(everything(), names_to = "parameter") |>
  separate(parameter, c("parameter", "bound")) |>
  pivot_wider(names_from = bound) |>
  flextable()
```

```{r}
#| message: false
#| warning: false
data_rec <- data |>
  mutate(sitenum = as.numeric(as.factor(Site))) |>
  subset(Treatment == "Logging" & trec > 0) |>
  # problem with first years in SUAS: excessive prod. because of  DBH increases
  filter(!(Site == "SUAS" & trec < 3 & variable == "prod")) |>
  filter(!(Site == "Tapajos Km67" &
             Plot %in% c(1:12, paste0(1:12, "E")) &
             Year >= 2014)) |> # remove second logging event in Tapajos
  pivot_wider(names_from = "variable", values_from = "value") |>
  subset(!is.na(prod) & !is.na(mort)) |>
  mutate(plotnum = as.numeric(as.factor(paste(Site, plot_new))))
data_old <- data |>
  mutate(sitenum = as.numeric(as.factor(Site))) |>
  subset(Treatment == "Control" | trec < 0) |>
  pivot_wider(names_from = "variable", values_from = "value") |>
  subset(!is.na(prod) & !is.na(mort))
```

```{r}
#| message: false
#| warning: false
#| eval: false
ind_rec <- data_rec |>
  select(Site, Plot, sitenum, plotnum) |>
  unique() |>
  arrange(plotnum)
data_t0 <- data_rec |>
  group_by(plotnum) |>
  summarise(t0 = trec[which.min(Year)], y0 = stocks[which.min(Year)])
mdata_all <- list(
  n_equ = nrow(data_old),
  n_log = nrow(data_rec),
  s = max(data_old$sitenum),
  p_log = max(data_rec$plotnum),
  y_equ = data_old$stocks,
  in_equ = data_old$prod,
  out_equ = data_old$mort,
  y_log = data_rec$stocks,
  in_log = data_rec$prod,
  out_log = data_rec$mort,
  time = data_rec$trec,
  dt = data_rec$tdiff,
  t0 = data_t0$t0,
  y0 = data_t0$y0,
  site_equ = data_old$sitenum,
  site_log = data_rec$sitenum,
  plot_log = data_rec$plotnum,
  site_plot = ind_rec$sitenum,
  theta_bounds = c(bounds$theta_min, bounds$theta_max),
  lambda_bounds = c(bounds$lambda_min, bounds$lambda_max),
  alpha_bounds = c(bounds$alpha_min, bounds$alpha_max),
  delta_bounds = c(bounds$delta_min, bounds$delta_max),
  tau_bounds = c(bounds$tau_min, bounds$tau_max),
  k_bounds = c(bounds$k_min, bounds$k_max),
  omega_bounds = c(bounds$omega_min, bounds$omega_max)
)
sample_model("flux_all_simpl_Ksite_nohyper", mdata_all)
```

The resulting Stan fit summary is presented below.

```{r}
#| message: false
#| warning: false
fit_flux_all_ks_nh <- as_cmdstan_fit(
  list.files("chains/flux_all_simpl_Ksite_nohyper/", full.names = TRUE)
)
fit_flux_all_ks_nh
mcmc_trace(fit_flux_all_ks_nh$draws(variables = c("lp__")))
```

Rhat superior to 1.1:

```{r}
rhat_all <- summarise_draws(fit_flux_all_ks_nh$draws(), "rhat")
rhat_all |>
  filter(rhat > 1.1) |>
  arrange(-rhat)
```

The goodness of fit is satisfactory.

```{r}
#| warning: false
#| fig-cap: "Predicted vs observed values."
#| fig-width: 8
#| fig-height: 3

pred_data <- data_rec |>
  select(-plotnum) |>
  mutate(
    prod = fit_flux_all_ks_nh$summary(c("mu_in_log"), median)$median,
    mort = fit_flux_all_ks_nh$summary(c("mu_out_log"), median)$median,
    val = "pred",
    type = "rec"
  ) |>
  rbind(mutate(
    data_old,
    prod = fit_flux_all_ks_nh$summary(c("theta_s"), median)$median[data_old$sitenum] * # nolint
      fit_flux_all_ks_nh$summary(c("omega_s"), median)$median[data_old$sitenum], # nolint
    mort = fit_flux_all_ks_nh$summary(c("theta_s"), median)$median[data_old$sitenum] * # nolint
      fit_flux_all_ks_nh$summary(c("omega_s"), median)$median[data_old$sitenum], # nolint
    val = "pred",
    type = "equ"
  ))

obs_data <- data_rec |>
  mutate(
    val = "obs",
    type = "rec"
  ) |>
  select(-plotnum) |>
  rbind(mutate(
    data_old,
    val = "obs",
    type = "equ"
  ))

pov_data <- rbind(pred_data, obs_data) |>
  pivot_longer(cols = c(prod, mort), names_to = "variable") |>
  pivot_wider(names_from = val, values_from = value)

df_rmse <- pov_data |>
  group_by(variable) |>
  summarise(rmse = sqrt(mean((pred - obs)^2)))

labs <- c(prod = "AGB productivity", mort = "AGB mortality")

pov_data |>
  ggplot(aes(pred, obs)) +
  geom_point(alpha = 0.25, aes(col = Site, shape = type)) +
  geom_abline(col = "red", linetype = "dashed") +
  geom_text(
    data = df_rmse,
    aes(
      x = -Inf, y = Inf, hjust = -0.1, vjust = 1,
      label = paste("RMSE =", round(rmse, 2))
    )
  ) +
  theme_bw() +
  xlab("Predicted") +
  ylab("Observed") +
  facet_wrap(~variable, scales = "free", labeller = labeller(variable = labs))
```

### Parameters (k and tau)

```{r}
fit_flux_all_ks_nh$summary(c("k_s")) |>
  separate(variable, c("variable", "sitenum"), sep = "\\[") |>
  mutate(sitenum = as.integer(gsub("]", "", sitenum))) |>
  inner_join(unique(data_rec[, c("Site", "sitenum")])) |>
  select(-sitenum) |>
  relocate(variable, Site)
```

```{r}
fit_flux_all_ks_nh$summary(c("mu_tau", "sigma_tau", "tau_s")) |>
  separate(variable, c("variable", "sitenum"), sep = "\\[") |>
  mutate(sitenum = as.integer(gsub("]", "", sitenum))) |>
  left_join(unique(data_rec[, c("Site", "sitenum")])) |>
  select(-sitenum) |>
  relocate(variable, Site)
```

### Visualising the trajectories

The predicted flux and stocks trajectories look consistent with the data.

```{r}
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 8
#| fig-cap: "Predicted trajectories."
pov_data |>
  subset(variable == "prod") |>
  mutate(Plot = plot_new) |>
  graph_traj(expression("AGB productivity [" ~ Mg ~ ha^-1 ~ yr^-1 ~ "]"))
pov_data |>
  subset(variable == "mort") |>
  mutate(Plot = plot_new) |>
  graph_traj(expression("AGB mortality [" ~ Mg ~ ha^-1 ~ yr^-1 ~ "]"))
```

### Predicted AGB recovery

The predicted AGB recovery can be estimated as the integral of the difference between productivity and mortality over the entire 

```{r}
data_t0 <- data_rec |>
  group_by(plotnum) |>
  summarise(t0 = trec[which.min(Year)], y0 = stocks[which.min(Year)])
theta <- fit_flux_all_ks_nh$summary(c("theta_s")) |>
  separate(variable, c("variable", "sitenum"), sep = "\\[") |>
  mutate(sitenum = as.integer(gsub("]", "", sitenum))) |>
  inner_join(unique(data_rec[, c("Site", "sitenum")])) |>
  select(Site, median, q5, q95) |>
  rename(theta_median = median, theta_q5 = q5, theta_q95 = q95)
agb_rec <- fit_flux_all_ks_nh$summary(c("diff_log")) |>
  separate(variable, c("variable", "plotnum"), sep = "\\[") |>
  mutate(plotnum = as.integer(gsub("]", "", plotnum))) |>
  inner_join(unique(data_rec[, c("Site", "Plot", "plotnum")])) |>
  left_join(data_t0) |>
  mutate(rec_median = y0 + median, rec_q5 = q5 + y0, rec_q95 = q95 + y0) |>
  select(Site, Plot, rec_median, rec_q5, rec_q95) |>
  left_join(theta)
ggplot(agb_rec, aes(theta_median, rec_median, col = Site)) +
  geom_abline(intercept = 0, slope = 1, lty = 2) +
  geom_pointrange(aes(ymin = rec_q5, ymax = rec_q95)) +
  geom_errorbar(aes(xmin = theta_q5, xmax = theta_q95), orientation = "y") +
  theme_minimal() +
  labs(
    x = "Estimated AGB of unlogged plots (parameter theta)",
    y = "Estimated AGB of logged plots after recovery"
  )
```