```{r setup}
#| message: false
#| warning: false
#| include: false

library(tidyverse)
library(cmdstanr)
library(bayesplot)
library(patchwork)
library(flextable)
library(posterior)
list.files("r", full.names = TRUE, pattern = ".R") |>
  lapply(source)
```

# Model calibration {.unnumbered}

## Data

```{r}
#| message: false
#| warning: false
plot_info <- read.csv("data/raw_data/bioforest-plot-information.csv") |>
  left_join(read.csv("data/raw_data/plot_correspondance.csv")) |>
  select(site, plot = new_plot, Year_of_harvest, Treatment) |>
  unique()

data <- read.csv("data/derived_data/aggregated_data_v10.csv") |>
  left_join(plot_info, by = join_by(Site == site, Plot == plot)) |>
  filter(grepl("^agb", variable) & !grepl("]$", variable)) |>
  pivot_wider(names_from = "variable", values_from = "value") |>
  mutate(prod = agb_growth + agb_recr, mort = agb_mort, stocks = agb) |>
  select(!contains("agb")) |>
  group_by(Site, plot_new = Plot) |>
  arrange(Year) |>
  mutate(tdiff = c(0, diff(Year))) |>
  ungroup() |>
  pivot_longer(cols = c("stocks", "prod", "mort"), names_to = "variable") |>
  mutate(Year_of_harvest = ifelse(Site == "Mbaiki" & Treatment == "Logging",
    1985, Year_of_harvest
  )) |>
  mutate(Year_of_harvest = ifelse(Site == "Paracou" & Treatment == "Logging",
    1987, Year_of_harvest
  )) |>
  mutate(trec = Year - Year_of_harvest) |>
  filter(!(Site == "Tapajos Km67" &
             grepl("loggedtwice", Plot) & Year > 2014)) |>
  filter(!(Site == "Tapajos Km67" & grepl("later", Plot)))
```

For now, we only calibrate the model with sites that meet all the criteria listed [here](https://bioforest-project.github.io/demography/21_flux_data.html), and that have at least 10 censuses.

```{r}
#| message: false
#| warning: false
sites <- read.csv("data/derived_data/sites_to_keep.csv") |>
  filter(ncensus >= 10)
data <- data |>
  filter(Site %in% sites$Site)
```

## Model

We used the following parameter bounds when calibrating the model.

```{r}
#| message: false
#| warning: false
bounds <- data.frame(
  "theta_min" = 50,
  "theta_max" = 1000,
  "lambda_min" = 1e-2,
  "lambda_max" = 5,
  "alpha_min" = 0,
  "alpha_max" = 200,
  "delta_min" = 0,
  "delta_max" = 20,
  "tau_min" = 2,
  "tau_max" = 30,
  "k_min" = 0.1,
  "k_max" = 5,
  "omega_min" = 1e-3,
  "omega_max" = 0.5
)
bounds |>
  pivot_longer(everything(), names_to = "parameter") |>
  separate(parameter, c("parameter", "bound")) |>
  pivot_wider(names_from = bound) |>
  flextable()
```

```{r}
#| message: false
#| warning: false
data_rec <- data |>
  mutate(sitenum = as.numeric(as.factor(Site))) |>
  subset(Treatment == "Logging" & trec > 0) |>
  # problem with first years in SUAS: excessive prod. because of  DBH increases
  filter(!(Site == "SUAS" & trec < 3 & variable == "prod")) |>
  filter(!(Site == "Tapajos Km67" &
             Plot %in% c(1:12, paste0(1:12, "E")) &
             Year >= 2014)) |> # remove second logging event in Tapajos
  pivot_wider(names_from = "variable", values_from = "value") |>
  subset(!is.na(prod) & !is.na(mort)) |>
  mutate(plotnum = as.numeric(as.factor(paste(Site, plot_new))))
data_old <- data |>
  mutate(sitenum = as.numeric(as.factor(Site))) |>
  subset(Treatment == "Control" | trec < 0) |>
  pivot_wider(names_from = "variable", values_from = "value") |>
  subset(!is.na(prod) & !is.na(mort))
```

```{r}
#| message: false
#| warning: false
#| eval: false
ind_rec <- data_rec |>
  select(Site, Plot, sitenum, plotnum) |>
  unique() |>
  arrange(plotnum)
mdata_all <- list(
  n_equ = nrow(data_old),
  n_log = nrow(data_rec),
  s = max(data_old$sitenum),
  p_log = max(data_rec$plotnum),
  y_equ = data_old$stocks,
  in_equ = data_old$prod,
  out_equ = data_old$mort,
  y_log = data_rec$stocks,
  in_log = data_rec$prod,
  out_log = data_rec$mort,
  time = data_rec$trec,
  dt = data_rec$tdiff,
  site_equ = data_old$sitenum,
  site_log = data_rec$sitenum,
  plot_log = data_rec$plotnum,
  site_plot = ind_rec$sitenum,
  theta_bounds = c(bounds$theta_min, bounds$theta_max),
  lambda_bounds = c(bounds$lambda_min, bounds$lambda_max),
  alpha_bounds = c(bounds$alpha_min, bounds$alpha_max),
  delta_bounds = c(bounds$delta_min, bounds$delta_max),
  tau_bounds = c(bounds$tau_min, bounds$tau_max),
  k_bounds = c(bounds$k_min, bounds$k_max),
  omega_bounds = c(bounds$omega_min, bounds$omega_max),
  y0_bounds = c(bounds$y0_min, bounds$y0_max)
)
sample_model("flux_all_simpl_Ksite_nohyper", mdata_all)
```

The resulting Stan fit summary is presented below.

```{r}
#| message: false
#| warning: false
fit_flux_all <- as_cmdstan_fit(
  list.files("chains/flux_all_simpl_Ksite_nohyper/", full.names = TRUE)
)
fit_flux_all
mcmc_trace(fit_flux_all$draws(variables = c("lp__")))
```

Rhat superior to 1.2:

```{r}
rhat_all <- summarise_draws(fit_flux_all$draws(), "rhat")
rhat_all |>
  filter(rhat > 1.2) |>
  arrange(-rhat) |>
  head()
```

The goodness of fit is satisfactory.

```{r}
#| warning: false
#| fig-cap: "Predicted vs observed values."
#| fig-width: 8
#| fig-height: 3

pred_data <- data_rec |>
  select(-plotnum) |>
  mutate(
    prod = fit_flux_all$summary(c("mu_in_log"), median)$median,
    mort = fit_flux_all$summary(c("mu_out_log"), median)$median,
    val = "pred",
    type = "rec"
  ) |>
  rbind(mutate(
    data_old,
    prod = fit_flux_all$summary(c("theta_s"), median)$median[data_old$sitenum] * # nolint
      fit_flux_all$summary(c("omega_s"), median)$median[data_old$sitenum], # nolint
    mort = fit_flux_all$summary(c("theta_s"), median)$median[data_old$sitenum] * # nolint
      fit_flux_all$summary(c("omega_s"), median)$median[data_old$sitenum], # nolint
    val = "pred",
    type = "equ"
  ))

obs_data <- data_rec |>
  mutate(
    val = "obs",
    type = "rec"
  ) |>
  select(-plotnum) |>
  rbind(mutate(
    data_old,
    val = "obs",
    type = "equ"
  ))

pov_data <- rbind(pred_data, obs_data) |>
  pivot_longer(cols = c(prod, mort), names_to = "variable") |>
  pivot_wider(names_from = val, values_from = value)

df_rmse <- pov_data |>
  group_by(variable) |>
  summarise(rmse = sqrt(mean((pred - obs)^2)))

labs <- c(prod = "AGB productivity", mort = "AGB mortality")

pov_data |>
  ggplot(aes(pred, obs)) +
  geom_point(alpha = 0.25, aes(col = Site, shape = type)) +
  geom_abline(col = "red", linetype = "dashed") +
  geom_text(
    data = df_rmse,
    aes(
      x = -Inf, y = Inf, hjust = -0.1, vjust = 1,
      label = paste("RMSE =", round(rmse, 2))
    )
  ) +
  theme_bw() +
  xlab("Predicted") +
  ylab("Observed") +
  facet_wrap(~variable, scales = "free", labeller = labeller(variable = labs))
```

### Parameters (k and tau)

```{r}
fit_flux_all$summary(c("k_s")) |>
  separate(variable, c("variable", "sitenum"), sep = "\\[") |>
  mutate(sitenum = as.integer(gsub("]", "", sitenum))) |>
  inner_join(unique(data_rec[, c("Site", "sitenum")])) |>
  select(-sitenum) |>
  select(variable, Site, mean, median, sd, q5, q95, rhat)
```

```{r}
fit_flux_all$summary(c("mu_tau", "sigma_tau", "tau_s")) |>
  separate(variable, c("variable", "sitenum"), sep = "\\[") |>
  mutate(sitenum = as.integer(gsub("]", "", sitenum))) |>
  left_join(unique(data_rec[, c("Site", "sitenum")])) |>
  select(-sitenum) |>
  select(variable, Site, mean, median, sd, q5, q95, rhat)
```

### Visualising the trajectories

The predicted flux and stocks trajectories look consistent with the data.

```{r}
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 8
#| fig-cap: "Predicted trajectories."
pov_data |>
  subset(variable == "prod") |>
  mutate(Plot = plot_new) |>
  graph_traj(expression("AGB productivity [" ~ Mg ~ ha^-1 ~ yr^-1 ~ "]"))
pov_data |>
  subset(variable == "mort") |>
  mutate(Plot = plot_new) |>
  graph_traj(expression("AGB mortality [" ~ Mg ~ ha^-1 ~ yr^-1 ~ "]"))
```

### Predicted AGB recovery

The predicted AGB recovered at the end of the trajectory is estimated as the sum of the AGB at the start of the trajectory and the integral of the productivity minus the mortality fluxes between the start of the recovery period and $t = +\infty$.

```{r}
#| message: false
#| warning: false
#| eval: false
stpp_integral <- function(t0, delta, tau, k) {
  u0 <- k * t0 / tau
  log_i <- log(delta) + k + log(tau) - (k + 1) * log(k) +
    lgamma(k + 1) + pgamma(u0, k + 1, lower.tail = FALSE, log.p = TRUE)
  exp(log_i)
}

data_t0 <- data_rec |>
  group_by(plotnum) |>
  summarise(t0 = trec[which.min(Year)], agb0 = stocks[which.min(Year)])
site_plot_num <- data |>
  filter(Treatment == "Logging" & trec >= 0) |>
  select(Site, plot_new) |>
  mutate(sitenum = as.numeric(as.factor(Site))) |>
  mutate(plotnum = as.numeric(as.factor(paste(Site, plot_new)))) |>
  unique()

alpha <- extract_draws(fit_flux_all, "alpha_p") |>
  mutate(plotnum = as.numeric(gsub("alpha_p|\\[|\\]", "", variable))) |>
  select(-variable) |>
  rename(alpha = value)
lambda <- extract_draws(fit_flux_all, "lambda_s") |>
  mutate(sitenum = as.numeric(gsub("lambda_s|\\[|\\]", "", variable))) |>
  select(-variable) |>
  rename(lambda = value)
tau <- extract_draws(fit_flux_all, "tau_s") |>
  mutate(sitenum = as.numeric(gsub("tau_s|\\[|\\]", "", variable))) |>
  select(-variable) |>
  rename(tau = value)
kappa <- extract_draws(fit_flux_all, "k_s") |>
  mutate(sitenum = as.numeric(gsub("k_s|\\[|\\]", "", variable))) |>
  rename(k = value) |>
  select(-variable)
delta <- extract_draws(fit_flux_all, "delta_p") |>
  mutate(plotnum = as.numeric(gsub("delta_p|\\[|\\]", "", variable))) |>
  select(-variable) |>
  rename(delta = value)
theta <- extract_draws(fit_flux_all, "theta_s") |>
  mutate(sitenum = as.numeric(gsub("theta_s|\\[|\\]", "", variable))) |>
  rename(theta = value) |>
  select(-variable)
omega <- extract_draws(fit_flux_all, "omega_s") |>
  mutate(sitenum = as.numeric(gsub("omega_s|\\[|\\]", "", variable))) |>
  rename(omega = value) |>
  select(-variable)
t90_p <- inner_join(tau, kappa) |>
  group_by(Site, iter) |>
  summarise(t90p = optimize(
    function(t) {
      abs(
        integrate(function(x) (exp(1 - x / tau) * x / tau)^k, 0, t)$value -
          0.9 * stpp_integral(0, 1, tau, k)
      )
    }, c(0, 100)
  )$minimum)

pars <- inner_join(site_plot_num, delta) |>
  inner_join(alpha) |>
  inner_join(tau) |>
  inner_join(kappa) |>
  inner_join(lambda) |>
  inner_join(theta) |>
  inner_join(omega) |>
  inner_join(t90_p) |>
  inner_join(data_t0) |>
  mutate(rec = agb0 + stpp_integral(t0, delta, tau, k) -
           alpha / lambda * exp(-lambda * t0)) |>
  mutate(stpp = stpp_integral(0, delta, tau, k)) |>
  mutate(stpm = alpha / lambda) |>
  mutate(t90m = log(10) / lambda) |>
  select(!contains(c("agb0", "t0", "num"))) |>
  pivot_longer(names_to = "variable", cols = !contains(c("ite", "lot")))

pars_plot <- pars |>
  filter(variable %in% c("alpha", "delta", "rec", "stpp", "stpm")) |>
  group_by(Site, plot_new, variable) |>
  summarise(
    mean = mean(value),
    median = median(value),
    sd = sd(value),
    q5 = quantile(value, 0.05),
    q95 = quantile(value, 0.95)
  )
pars_site <- pars |>
  filter(!variable %in% c("alpha", "delta", "rec", "stpp", "stpm")) |>
  group_by(Site, variable) |>
  summarise(
    mean = mean(value),
    median = median(value),
    sd = sd(value),
    q5 = quantile(value, 0.05),
    q95 = quantile(value, 0.95)
  )
bind_rows(pars_plot, pars_site) |>
  write.csv("data/derived_data/params_flux_all.csv", row.names = FALSE)
```

```{r}
#| message: false
#| warning: false
#| fig-width: 5
#| fig-height: 4
#| fig-cap: "Predicted AGB at the end of the recovery trajectories (as the integral of fluxes + the initial AGB) vs the equilibrium biomass parameter theta." #nolint
theta <- read.csv("data/derived_data/params_flux_all.csv") |>
  filter(variable == "theta") |>
  select(Site, median, variable, q5, q95) |>
  pivot_wider(names_from = variable, values_from = c(median, q5, q95))
read.csv("data/derived_data/params_flux_all.csv") |>
  filter(variable == "rec") |>
  select(Site, plot_new, median, variable, q5, q95) |>
  pivot_wider(names_from = variable, values_from = c(median, q5, q95)) |>
  inner_join(theta) |>
  ggplot(aes(median_theta, median_rec, col = Site)) +
  geom_abline(intercept = 0, slope = 1, lty = 2) +
  geom_pointrange(aes(ymin = q5_rec, ymax = q95_rec)) +
  geom_errorbar(aes(xmin = q5_theta, xmax = q95_theta), orientation = "y") +
  theme_minimal() +
  labs(
    x = "Estimated AGB of unlogged plots (parameter theta)",
    y = "Estimated AGB of logged plots after recovery"
  )
```
