```{r setup}
#| message: false
#| warning: false
library(tidyverse)
```

## Stocks - with constant mortality {.unnumbered}

Here we define productivity with the general recovery model, and we hypothesize that mortality is a constant:

$$ prod(t) = \theta_0 + (\theta_{\infty} -\theta_0)\cdot\left(1 - exp(-\lambda\cdot t)+\delta \cdot \left(\frac{t}{\tau} \cdot exp\left(1 - \frac{t}{\tau}\right)\right)^2\right)$$

$$ba'(t) = prod(t) - k$$ We need $lim_{\infty}(prod) = lim_{\infty}(mort)$ to have a stable asymptotic basal area. We therefore have:

$k = \theta_{\infty}$

We can therefore write :

$$ba(t) =  ba(0) + \int_{t=0}^{\infty}prod(x) dx - \theta_{\infty}$$

After integrating the productivity equation (see Appendix "Integration of STP" for the integration ), we get the following equation for basal area: $$ba(t) = ba(0) + (\theta_{\infty} - \theta_0)\left(\frac{\delta \cdot e^{2} \cdot\tau}{4} - \delta \cdot e^{(1-\frac{t}{\tau})2} \left(\frac{t^2}{2\tau}+\frac{t}{2}+\frac{\tau}{4}\right) - \frac{1}{\lambda} (1-e^{-\lambda\cdot t})\right) $$

The figure below illustrates the resulting stocks and fluxes with parameters set to:

-   $\tau = 10$
-   $\lambda = 0.05$
-   $\delta = 1$
-   $\theta_{\infty} = 0.5$
-   $\theta_0 = 0.4$
-   $ba_0 = 10$

```{r, fig.width = 6, fig.height = 6}
thetainf <- 0.5
theta0 <- 0.4
ba0 <- 10
lambda <- 0.05

ltp <- function(x, lambda = 0.05) 1 - exp(-lambda * x)
stp <- function(time, delta = 1, tau = 10) {
  delta * (time / tau * exp(1 - time / tau))^2
}
int_stp <- function(time, delta = 1, tau = 10) {
  delta * exp(2) * tau / 4 - delta * exp((1 - time / tau) * 2) *
    (time^2 / (2 * tau) + time / 2 + tau / 4)
}

facet_labs <- c(
  "ba" = "Basal area [m2/ha]",
  "prod" = "Productivity [m2/ha/yr]",
  "mort" = "Mortality [m2/ha/yr]"
)

data.frame(t = seq(0, 100, length.out = 200)) |>
  mutate(
    prod = theta0 + (thetainf - theta0) * (stp(t) + ltp(t)),
    mort = thetainf,
    ba = ba0 + (thetainf - theta0) * (int_stp(t) - 1 / lambda *
                                        (1 - exp(-lambda * t)))
  ) |>
  pivot_longer(-t) |>
  ggplot(aes(t, value)) +
  geom_line() +
  facet_wrap(~name,
    scales = "free", nrow = 3,
    labeller = labeller(name = facet_labs),
    strip.position = "left"
  ) +
  labs(x = "Recovery time [yr]", y = NULL) +
  theme_minimal() +
  theme(strip.placement = "outside")
```
