```{r setup}
#| message: false
#| warning: false
#| include: false

library(tidyverse)
library(cmdstanr)
list.files("r", full.names = TRUE, pattern = ".R") |>
  lapply(source)
```

Done for the model with a fitted kurtosis per site without hyperlaw.

# Explanatory variables {.unnumbered}

We chose the following summary statistics:

- the turnover rate $\omega$;

- the total over-productivity and over-mortality are estimated as the area under the curve of the short-term process:

- $\frac{\delta e^2 \tau}{4}$ for productivity;

- $\frac{\alpha}{\lambda}$ for mortality;

- the time taken to reach 95% of the total over-productivity and mortality (i.e. the duration of the change in fluxes):

- the numerical solution to the equation: $e^{-\frac{2}{\tau} t_{95}}(\frac{t_{95}^2}{2\tau}+\frac{t_{95}}{2}+\frac{\tau}{4})=\frac{\tau}{80}$ for productivity;

- $t_{95} = \frac{ln(20)}{\lambda}$ for mortality.

```{r}
#| message: false
#| warning: false
plot_info <- read.csv("data/raw_data/bioforest-plot-information.csv") |>
  select(site, plot, Year_of_harvest, Treatment)
sites <- read.csv("data/derived_data/sites_to_keep.csv") |>
  filter(ncensus >= 10)
data <- read.csv("data/derived_data/aggregated_data_v9.csv") |>
  mutate(plot_new = Plot) |>
  separate(Plot, "Plot", "_") |>
  left_join(plot_info, by = join_by(Site == site, Plot == plot)) |>
  filter(Site %in% sites$Site) |>
  filter(grepl("^agb", variable)) |>
  pivot_wider(names_from = "variable", values_from = "value") |>
  mutate(prod = agb_growth + agb_recr, mort = agb_mort, stocks = agb) |>
  select(!contains("agb")) |>
  pivot_longer(cols = c("stocks", "prod", "mort"), names_to = "variable") |>
  mutate(Year_of_harvest = ifelse(Site == "Mbaiki" & Treatment == "Logging",
    1985, Year_of_harvest
  )) |>
  mutate(trec = Year - Year_of_harvest)
```

## Effect of environmental variables

```{r}
#| message: false
#| warning: false
environ <- readr::read_tsv("data/raw_data/environment.tsv")
params <- read.csv("data/derived_data/params_flux_all.csv")
```

```{r}
#| message: false
#| warning: false
environ |>
  group_by(site) |>
  summarise(pr = mean(dsi)) |>
  inner_join(filter(params, variable == "omega"), by = join_by(site == Site)) |>
  ggplot(aes(pr, median, col = site)) +
  geom_pointrange(aes(ymin = q5, ymax = q95)) +
  labs(x = "Dry season intensity", y = "AGB turnover rate") +
  theme_minimal()
```

```{r}
#| message: false
#| warning: false
environ |>
  group_by(site) |>
  summarise(pr = mean(pr)) |>
  inner_join(filter(params, variable == "t90p"), by = join_by(site == Site)) |>
  ggplot(aes(pr, median, col = site)) +
  geom_pointrange(aes(ymin = q5, ymax = q95)) +
  labs(
    x = "Mean annual precipitation [ mm ]",
    y = "AGB overproductivity duration [ yr ]"
  ) +
  theme_minimal()
```

## Effect of logging intensity

```{r}
#| message: false
#| warning: false
read.csv("data/derived_data/aggregated_data_v10.csv") |>
  filter(Site %in% sites$Site & variable == "ba") |>
  mutate(plot_new = Plot) |>
  separate(Plot, "Plot", "_") |>
  left_join(plot_info, by = join_by(Site == site, Plot == plot)) |>
  group_by(Site) |>
  mutate(ba_control = mean(value[Treatment == "Control"])) |>
  group_by(Site, plot_new, Treatment) |>
  summarise(
    ba_0 = ifelse(Year_of_harvest <= min(Year),
      ba_control, value[Year == min(Year)]
    ),
    ba_min = min(value[Year > Year_of_harvest & Year <= Year_of_harvest + 5])
  ) |>
  mutate(intensity = (ba_0 - ba_min) / ba_0) |>
  inner_join(filter(params, variable == "stpp")) |>
  ggplot(aes(intensity * 100, median, col = Site)) +
  geom_pointrange(aes(ymin = q5, ymax = q95)) +
  labs(
    x = "Logging intensity [% BA loss]",
    y = "AGB total overproductivity [Mg/ha]"
  ) +
  theme_minimal()
```

```{r}
#| message: false
#| warning: false
read.csv("data/derived_data/aggregated_data_v10.csv") |>
  filter(Site %in% sites$Site & variable == "ba") |>
  mutate(plot_new = Plot) |>
  separate(Plot, "Plot", "_") |>
  left_join(plot_info, by = join_by(Site == site, Plot == plot)) |>
  group_by(Site) |>
  mutate(ba_control = mean(value[Treatment == "Control"])) |>
  group_by(Site, plot_new, Treatment) |>
  summarise(
    ba_0 = ifelse(Year_of_harvest <= min(Year),
      ba_control, value[Year == min(Year)]
    ),
    ba_min = min(value[Year > Year_of_harvest & Year <= Year_of_harvest + 5])
  ) |>
  mutate(intensity = (ba_0 - ba_min) / ba_0) |>
  inner_join(filter(params, variable == "stpm")) |>
  ggplot(aes(intensity * 100, median, col = Site)) +
  geom_pointrange(aes(ymin = q5, ymax = q95)) +
  labs(
    x = "Logging intensity [% BA loss]",
    y = "AGB total overmortality [Mg/ha]"
  ) +
  theme_minimal()
```
