```{r setup}
#| message: false
#| warning: false
#| include: false

library(tidyverse)
list.files("r", full.names = TRUE, pattern = ".R") %>%
  lapply(source)
```

# Explanatory variables {.unnumbered}

We chose the following summary statistics:

-   the turnover rate $\omega$;

-   the total over-productivity and over-mortality are estimated as the area under the curve of the short-term process:

-   $\frac{\delta e^2 \tau}{4}$ for productivity;

-   $\frac{\alpha}{\lambda}$ for mortality;

-   the time taken to reach 95% of the total over-productivity and mortality (i.e. the duration of the change in fluxes):

-   the numerical solution to the equation: $e^{-\frac{2}{\tau} t_{95}}(\frac{t_{95}^2}{2\tau}+\frac{t_{95}}{2}+\frac{\tau}{4})=\frac{\tau}{80}$ for productivity;

-   $t_{95} = \frac{ln(20)}{\lambda}$ for mortality.

```{r}
#| message: false
#| warning: false
plot_info <- read.csv("data/raw_data/bioforest-plot-information.csv") |>
  select(site, plot, Year_of_harvest, Treatment)
sites <- read.csv("data/derived_data/sites_to_keep.csv") |>
  filter(ncensus >= 10)
data <- read.csv("data/derived_data/aggregated_data_v9.csv") |>
  mutate(plot_new = Plot) |>
  separate(Plot, "Plot", "_") |>
  left_join(plot_info, by = join_by(Site == site, Plot == plot)) |>
  filter(Site %in% sites$Site) |>
  filter(grepl("^agb", variable)) |>
  pivot_wider(names_from = "variable", values_from = "value") |>
  mutate(prod = agb_growth + agb_recr, mort = agb_mort, stocks = agb) |>
  select(!contains("agb")) |>
  pivot_longer(cols = c("stocks", "prod", "mort"), names_to = "variable") |>
  mutate(Year_of_harvest = ifelse(Site == "Mbaiki" & Treatment == "Logging",
    1985, Year_of_harvest
  )) |>
  mutate(trec = Year - Year_of_harvest)
```

```{r}
#| message: false
#| warning: false
#| eval: false
fit_flux_all <-
  as_cmdstan_fit(list.files("chains/flux_all/", full.names = TRUE))
site_plot_num <- data |>
  filter(Treatment == "Logging" & trec > 3) |>
  select(Site, plot_new) |>
  mutate(sitenum = as.numeric(as.factor(Site))) |>
  mutate(plotnum = as.numeric(as.factor(paste(Site, plot_new)))) |>
  unique()

alpha <- extract_draws(fit_flux_all, "alpha_p") |>
  separate(variable, into = c(NA, NA, "plotnum", NA), remove = FALSE) |>
  mutate(plotnum = as.numeric(plotnum)) |>
  rename(alpha = value) |>
  left_join(site_plot_num)
lambda <- extract_draws(fit_flux_all, "lambda_s") |>
  separate(variable, into = c(NA, NA, "sitenum", NA)) |>
  mutate(sitenum = as.numeric(sitenum)) |>
  rename(lambda = value) |>
  left_join(unique(select(site_plot_num, contains("ite"))))
tau <- extract_draws(fit_flux_all, "tau_s") |>
  separate(variable, into = c(NA, NA, "sitenum", NA)) |>
  mutate(sitenum = as.numeric(sitenum)) |>
  rename(tau = value) |>
  left_join(unique(select(site_plot_num, contains("ite"))))
delta <- extract_draws(fit_flux_all, "delta_p") |>
  separate(variable, into = c(NA, NA, "plotnum", NA), remove = FALSE) |>
  mutate(plotnum = as.numeric(plotnum)) |>
  rename(delta = value) |>
  left_join(site_plot_num)

int_p <- delta |>
  inner_join(tau, by = join_by(Site, sitenum, iter)) |>
  mutate(value = delta * exp(2) * tau / 4, variable = "int_p") |>
  group_by(Site, plot_new, variable) |>
  summarise(
    mean = mean(value),
    median = median(value),
    q5 = quantile(value, 0.05),
    q95 = quantile(value, 0.95)
  ) |>
  select(Site, plot_new, variable, mean, median, q5, q95)
int_m <- alpha |>
  inner_join(lambda, by = join_by(Site, sitenum, iter)) |>
  mutate(value = alpha / lambda, variable = "int_m") |>
  group_by(Site, plot_new, variable) |>
  summarise(
    mean = mean(value),
    median = median(value),
    q5 = quantile(value, 0.05),
    q95 = quantile(value, 0.95)
  ) |>
  select(Site, plot_new, variable, mean, median, q5, q95)

eq_t95_p <- function(x, tau) {
  abs(exp(-2 / tau * x) * (x^2 / (2 * tau) + x / 2 + tau / 4) - tau / 80)
}
t95_p <- tau |>
  group_by(Site, iter) |>
  mutate(value = optimize(function(x) eq_t95_p(x, tau), c(0, 100))$minimum) |>
  group_by(Site) |>
  summarise(
    mean = mean(value),
    median = median(value),
    q5 = quantile(value, 0.05),
    q95 = quantile(value, 0.95)
  ) |>
  mutate(variable = "t95_p") |>
  select(Site, variable, mean, median, q5, q95)
t95_m <- lambda |>
  mutate(value = log(20) / lambda) |>
  group_by(Site) |>
  summarise(
    mean = mean(value),
    median = median(value),
    q5 = quantile(value, 0.05),
    q95 = quantile(value, 0.95)
  ) |>
  mutate(variable = "t95_m") |>
  select(Site, variable, mean, median, q5, q95)

fit_flux_all$summary(variables = "omega_s") |>
  separate(variable, into = c("variable", NA, "sitenum", NA)) |>
  mutate(sitenum = as.numeric(sitenum)) |>
  left_join(unique(select(site_plot_num, contains("ite")))) |>
  select(Site, variable, mean, median, q5, q95) |>
  bind_rows(int_p) |>
  bind_rows(int_m) |>
  bind_rows(t95_p) |>
  bind_rows(t95_m) |>
  write.csv("data/derived_data/params_flux_all.csv", row.names = FALSE)
```

## Effect of environmental variables 

```{r}
#| message: false
#| warning: false
environ <- readr::read_tsv("data/raw_data/environment.tsv") |>
  mutate(site = gsub("Tapajós km 67", "Tapajos Km67", site))
params <- read.csv("data/derived_data/params_flux_all.csv")
```

```{r}
#| message: false
#| warning: false
environ |>
  group_by(site) |>
  summarise(pr = mean(pr)) |>
  inner_join(filter(params, variable == "omega"), by = join_by(site == Site)) |>
  ggplot(aes(pr, median, col = site)) +
  geom_pointrange(aes(ymin = q5, ymax = q95)) +
  labs(x = "Mean annual precipitation [ mm ]", y = "AGB turnover rate") +
  theme_minimal()
```

```{r}
#| message: false
#| warning: false
environ |>
  group_by(site) |>
  summarise(pr = mean(pr)) |>
  inner_join(filter(params, variable == "t95_p"), by = join_by(site == Site)) |>
  ggplot(aes(pr, median, col = site)) +
  geom_pointrange(aes(ymin = q5, ymax = q95)) +
  labs(
    x = "Mean annual precipitation [ mm ]",
    y = "AGB overproductivity duration [ yr ]"
  ) +
  theme_minimal()
```

## Effect of logging intensity

```{r}
#| message: false
#| warning: false
read.csv("data/derived_data/aggregated_data_v9.csv") |>
  filter(Site %in% sites$Site & variable == "ba") |>
  mutate(plot_new = Plot) |>
  separate(Plot, "Plot", "_") |>
  left_join(plot_info, by = join_by(Site == site, Plot == plot)) |>
  group_by(Site, plot_new, Treatment) |>
  summarise(
    ba_0 = value[Year == min(Year)],
    ba_min = min(value[Year > Year_of_harvest & Year <= Year_of_harvest + 5])
  ) |>
  mutate(intensity = (ba_0 - ba_min) / ba_0) |>
  inner_join(filter(params, variable == "int_p")) |>
  ggplot(aes(intensity * 100, median, col = Site)) +
  geom_pointrange(aes(ymin = q5, ymax = q95)) +
  labs(
    x = "Logging intensity [% BA loss]",
    y = "AGB total overproductivity [Mg/ha]"
  ) +
  theme_minimal()
```

