```{r setup}
#| message: false
#| warning: false
#| include: false
library(tidyverse)
list.files("r", full.names = TRUE, pattern = ".R") %>%
  lapply(source)
```

## Stocks - with proportional mortality {.unnumbered}

Here we define productivity with the general recovery model, and we hypothesize that mortality is proportional to basal area:

$$ prod(t) = \theta_0 + (\theta_{\infty} -\theta_0)\cdot\left(1 - exp(-\lambda\cdot t) \right) + \theta_{\infty}\cdot\delta \left(\frac{t}{\tau} \cdot exp\left(1 - \frac{t}{\tau}\right)\right)^2$$

$$ba'(t) = prod(t) - \omega \cdot ba(t)$$ and the initial condition:

$$ba(t=0) = BA_0$$

The solution to this equation is:

$$ba(t) = \frac{\theta_{\infty}}{\omega} + Kexp(-\omega t) + \frac{\lambda(\theta_{\infty}-\theta_0) }{1-\omega\lambda} \cdot exp(-\lambda\cdot t) + exp(1-\frac{t}{\tau})^2\frac{2\delta\theta_{\infty}}{\tau^2(2\omega-\tau)}\left(t^2 - t\frac{4}{(2\omega-\tau)} + \frac{8}{(2\omega-\tau)^2}\right)$$

with:

$$ K = ba_0 - \frac{\theta_{\infty}}{\omega} - \frac{\lambda(\theta_{\infty}-\theta_0) }{1-\omega\lambda} -\frac{16e^2\delta\theta_{\infty}}{\tau^2(2\omega-\tau)^3}$$ (see Appendix "Differential equation")

Below is a graph of the resulting fluxes and stocks with parameter values:

-   $\tau = 10$
-   $\lambda = 0.05$
-   $\delta = 0.2$
-   $\theta_{\infty} = 0.5$
-   $\theta_0 = 0.4$
-   $\omega = 0.02$
-   $ba_0 = 10$

```{r, fig.width = 6, fig.height = 6}
facet_labs <- c(
  "ba" = "Basal area [m2/ha]",
  "prod" = "Productivity [m2/ha/yr]",
  "mort" = "Mortality [m2/ha/yr]"
)

data.frame(t = seq(0, 200, 1e-2)) |>
  mutate(
    ba = int_stocks(t, tau = 5),
    prod = productivity(t, tau = 5),
    mort = mortality(t, tau = 5)
  ) |>
  pivot_longer(-t) |>
  ggplot(aes(t, value)) +
  geom_line() +
  scale_x_sqrt() +
  facet_wrap(~name,
    scales = "free", nrow = 3,
    labeller = labeller(name = facet_labs),
    strip.position = "left"
  ) +
  labs(x = "Recovery time [yr]", y = NULL) +
  theme_minimal() +
  theme(strip.placement = "outside")
```

To identify which parameters of the model have the strongest effect on the recovery rate of the basal area, we performed a sensitivity analysis by making each parameter vary while others remained constant, and estimated the time (in years) to recover 95% of basal area lost, estimated as the asymptotic basal area (equal to $\frac{\theta_{\infty}}{\omega}$) minus the post-logging basal area $BA_0$.

```{r}
# what parameters have the most effect on the recovery rate of ba?
pars <- data.frame(
  param = c(
    "tau", "lambda", "delta", "theta[infinity]",
    "theta[0]", "omega", "ba[0]"
  ),
  min = c(1, 1e-3, 1e-3, 0.1, 0.1, 0.001, 1e-3),
  max = c(50, 1, 10, 1, 1, 0.2, 20)
)

get_t95 <- function(tau, lambda, delta, thetainf, theta0, omega, ba0) {
  min(which(
    int_stocks(1:1e4, tau, lambda, delta, thetainf, theta0, omega, ba0) > 0.95 *
      (thetainf / omega - ba0) + ba0
  ))
}
pars |>
  group_by(param) |>
  reframe(value = exp(seq(log(min), log(max), length.out = 10))) |>
  cbind(data.frame(
    tau = 10, lambda = 0.05, delta = 0.2, thetainf = 0.5,
    theta0 = 0.4, omega = 0.02, ba0 = 10
  )) |>
  mutate(
    tau = ifelse(param == "tau", value, tau), ,
    lambda = ifelse(param == "lambda", value, lambda),
    delta = ifelse(param == "delta", value, delta),
    thetainf = ifelse(param == "theta[infinity]", value, thetainf),
    theta0 = ifelse(param == "theta[0]", value, theta0),
    omega = ifelse(param == "omega", value, omega),
    ba0 = ifelse(param == "ba[0]", value, ba0)
  ) |>
  # constraints: theta0 < thetainf and ba0 < bainf = thetainf/omega
  subset(theta0 < thetainf & ba0 < thetainf / omega) |>
  rowwise() |>
  mutate(t95 = get_t95(tau, lambda, delta, thetainf, theta0, omega, ba0)) |>
  ggplot(aes(value, t95)) +
  geom_line() +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap(~param,
    scales = "free",
    labeller = label_parsed,
    strip.position = "bottom"
  ) +
  labs(x = NULL, y = "t95 [yr]") +
  theme_minimal() +
  theme(strip.placement = "outside")
```
