```{r setup}
#| message: false
#| warning: false
#| include: false

library(tidyverse)
library(dplyr)
list.files("r", full.names = TRUE, pattern = ".R") %>%
  lapply(source)
```

# Composition change data {.unnumbered}

## Formula for the contribution of fluxes

The contribution of a given flux (recruitment, growth or mortality) at time $t$ can be written as:

$$CF_{F,t} = \sum_{i\in F_t}\frac{ba_{i,t}  (ft_{i} - CWM_{t-1})}{BA_{t}}-\sum_{i\in F_t}\frac{ba_{i,t-1} (ft_{i} - CWM_{t-1})}{BA_{t-1}}$$ where $F_t$ is the set of trees contributing to a given flux between $t-1$ and $t$ (recruitment: trees recruited at $t$; growth: trees that were measured and alive at $t$ and $t-1$; mortality: trees that died between $t-1$ and $t$).

$ba_{i,t}$ is the basal area of a tree $i$ at time $t$; $ft_{i}$ is the functional trait value of tree $i$; $BA_t$ is the basal area of the entire community at $t$; $CWM_t$ is the community weighted mean trait value at $t$.

We subtract $CWM_{t-1}$ to have negative contributions when the flux decreases the CWM between $t-1$ and $t$; and positive contributions that increase the $CWM$.

As written, the contributions are additive:

$$CWM_t= CWM_{t-1} + CF_{G,t} + CF_{R,t}+CF_{M,t}$$

## Data

```{r}
#| message: false
#| warning: false
plot_info <- read.csv("data/raw_data/bioforest-plot-information.csv") |>
  select(site, plot, Year_of_harvest, Treatment)
data <- read.csv("data/derived_data/aggregated_data_v9.csv") |>
  mutate(plot_new = Plot) |>
  separate(Plot, "Plot", "_") |>
  left_join(plot_info, by = join_by(Site == site, Plot == plot)) |>
  filter(grepl("cwmdiff", variable)) |>
  separate(variable, c("trait", NA, "flux"), sep = "_")
```

### Annual changes

::: panel-tabset

```{r}
#| results: asis
#| message: false
#| warning: false
#| fig.height: 7
#| fig.width: 10

sites <- c("Paracou", "Mbaiki", "SUAS")
labs_flux <- c(
  all = "Change in CWM",
  rba = "Contribution of recruitment",
  gba = "Contribution of growth",
  mba = "Contribution of mortality"
)
purrr::walk(
  sites,
  \(s) {
    cat("#### ", s, "\n\n")
    print(
      data |>
        filter(Site == s) |>
        ggplot(aes(x = Year, y = value, group = plot_new, col = Treatment)) +
        geom_line() +
        geom_point() +
        facet_grid(trait ~ flux,
          scales = "free",
          labeller = labeller(flux = labs_flux)
        ) +
        labs(y = NULL, x = NULL) +
        theme_minimal()
    )
    cat("\n\n")
  }
)
```

:::


### Cumulative changes


::: panel-tabset

```{r}
#| results: asis
#| message: false
#| warning: false
#| fig.height: 7
#| fig.width: 10

sites <- c("Paracou", "Mbaiki", "SUAS")
labs_flux <- c(
  all = "Change in CWM",
  rba = "Contribution of recruitment",
  gba = "Contribution of growth",
  mba = "Contribution of mortality"
)
purrr::walk(
  sites,
  \(s) {
    cat("#### ", s, "\n\n")
    print(
      data |>
        filter(Site == s) |>
        arrange(Site, plot_new, Year) |>
        group_by(Site, plot_new, trait, flux) |>
        mutate(value = cumsum(value)) |>
        ggplot(aes(x = Year, y = value, group = plot_new, col = Treatment)) +
        geom_line() +
        geom_point() +
        facet_grid(trait ~ flux,
          scales = "free",
          labeller = labeller(flux = labs_flux)
        ) +
        labs(y = NULL, x = NULL) +
        theme_minimal()
    )
    cat("\n\n")
  }
)
```

:::
