```{r setup}
#| message: false
#| warning: false
#| include: false

library(tidyverse)
library(cmdstanr)
library(bayesplot)
library(patchwork)
library(flextable)
library(posterior)
library(ggpubr)
list.files("r", full.names = TRUE, pattern = ".R") |>
  lapply(source)
```

# Testing different versions of the productivity model {.unnumbered}

Here we tested different versions of the demographic rates model, with some changes in the productivity function. 

## Data

```{r}
#| message: false
#| warning: false
plot_info <- read.csv("data/raw_data/bioforest-plot-information.csv") |>
  left_join(read.csv("data/raw_data/plot_correspondance.csv")) |>
  select(site, plot = new_plot, Year_of_harvest, Treatment) |>
  unique()

data <- read.csv("data/derived_data/aggregated_data_v10.csv") |>
  left_join(plot_info, by = join_by(Site == site, Plot == plot)) |>
  filter(grepl("^agb", variable) & !grepl("]$", variable)) |>
  pivot_wider(names_from = "variable", values_from = "value") |>
  mutate(prod = agb_growth + agb_recr, mort = agb_mort, stocks = agb) |>
  select(!contains("agb")) |>
  group_by(Site, plot_new = Plot) |>
  arrange(Year) |>
  mutate(tdiff = c(0, diff(Year))) |>
  ungroup() |>
  pivot_longer(cols = c("stocks", "prod", "mort"), names_to = "variable") |>
  mutate(Year_of_harvest = ifelse(Site == "Mbaiki" & Treatment == "Logging",
    1985, Year_of_harvest
  )) |>
  mutate(Year_of_harvest = ifelse(Site == "Paracou" & Treatment == "Logging",
    1987, Year_of_harvest
  )) |>
  mutate(trec = Year - Year_of_harvest) |>
  filter(!(Site == "Tapajos Km67" &
             grepl("loggedtwice", Plot) & Year > 2014)) |>
  filter(!(Site == "Tapajos Km67" & grepl("later", Plot)))

sites <- read.csv("data/derived_data/sites_to_keep.csv") |>
  filter(ncensus >= 10)
data <- data |>
  filter(Site %in% sites$Site)
```

We used the following parameter bounds when calibrating the model.

```{r}
#| message: false
#| warning: false
bounds <- data.frame(
  "theta_min" = 50,
  "theta_max" = 1000,
  "lambda_min" = 1e-2,
  "lambda_max" = 5,
  "alpha_min" = 0,
  "alpha_max" = 200,
  "delta_min" = 0,
  "delta_max" = 20,
  "tau_min" = 2,
  "tau_max" = 30,
  "k_min" = 0.1,
  "k_max" = 5,
  "omega_min" = 1e-3,
  "omega_max" = 0.5,
  "y0_min" = 10,
  "y0_max" = 1000
)
bounds |>
  pivot_longer(everything(), names_to = "parameter") |>
  separate(parameter, c("parameter", "bound")) |>
  pivot_wider(names_from = bound) |>
  flextable()
```

```{r}
#| message: false
#| warning: false
data_rec <- data |>
  mutate(sitenum = as.numeric(as.factor(Site))) |>
  subset(Treatment == "Logging" & trec > 0) |>
  # problem with first years in SUAS: excessive prod. because of  DBH increases
  filter(!(Site == "SUAS" & trec < 3 & variable == "prod")) |>
  filter(!(Site == "Tapajos Km67" &
             Plot %in% c(1:12, paste0(1:12, "E")) &
             Year >= 2014)) |> # remove second logging event in Tapajos
  pivot_wider(names_from = "variable", values_from = "value") |>
  subset(!is.na(prod) & !is.na(mort)) |>
  mutate(plotnum = as.numeric(as.factor(paste(Site, plot_new))))
data_old <- data |>
  mutate(sitenum = as.numeric(as.factor(Site))) |>
  subset(Treatment == "Control" | trec < 0) |>
  pivot_wider(names_from = "variable", values_from = "value") |>
  subset(!is.na(prod) & !is.na(mort))
```

## Different productivity model versions

### Constant kurtosis

In this version, the kurtosis parameter is set to $\kappa = 2$ and the productivity is therefore modelled as: 

$$ STPprod_{p,t} = \delta_p \cdot \left( \frac{t}{\tau_s} \cdot exp(1-\frac{t}{\tau_s}) \right)^{2} $$
```{r}
#| message: false
#| warning: false
#| eval: false
ind_rec <- data_rec |>
  select(Site, Plot, sitenum, plotnum) |>
  unique() |>
  arrange(plotnum)

mdata_all <- list(
  n_equ = nrow(data_old),
  n_log = nrow(data_rec),
  s = max(data_old$sitenum),
  p_log = max(data_rec$plotnum),
  y_equ = data_old$stocks,
  in_equ = data_old$prod,
  out_equ = data_old$mort,
  y_log = data_rec$stocks,
  in_log = data_rec$prod,
  out_log = data_rec$mort,
  time = data_rec$trec,
  dt = data_rec$tdiff,
  site_equ = data_old$sitenum,
  site_log = data_rec$sitenum,
  plot_log = data_rec$plotnum,
  site_plot = ind_rec$sitenum,
  theta_bounds = c(bounds$theta_min, bounds$theta_max),
  lambda_bounds = c(bounds$lambda_min, bounds$lambda_max),
  alpha_bounds = c(bounds$alpha_min, bounds$alpha_max),
  delta_bounds = c(bounds$delta_min, bounds$delta_max),
  tau_bounds = c(bounds$tau_min, bounds$tau_max),
  omega_bounds = c(bounds$omega_min, bounds$omega_max),
  y0_bounds = c(bounds$y0_min, bounds$y0_max)
)
sample_model("flux_all_simpl", mdata_all)
```

### Fitted kurtosis - common to all sites

In this version, the kurtosis parameter $\kappa$ is fitted, with one value common to all sites, and the productivity is therefore modelled as: 

$$ STPprod_{p,t} = \delta_p \cdot \left( \frac{t}{\tau_s} \cdot exp(1-\frac{t}{\tau_s}) \right)^{\kappa} $$

```{r}
#| message: false
#| warning: false
#| eval: false
ind_rec <- data_rec |>
  select(Site, Plot, sitenum, plotnum) |>
  unique() |>
  arrange(plotnum)

mdata_all <- list(
  n_equ = nrow(data_old),
  n_log = nrow(data_rec),
  s = max(data_old$sitenum),
  p_log = max(data_rec$plotnum),
  y_equ = data_old$stocks,
  in_equ = data_old$prod,
  out_equ = data_old$mort,
  y_log = data_rec$stocks,
  in_log = data_rec$prod,
  out_log = data_rec$mort,
  time = data_rec$trec,
  dt = data_rec$tdiff,
  site_equ = data_old$sitenum,
  site_log = data_rec$sitenum,
  plot_log = data_rec$plotnum,
  site_plot = ind_rec$sitenum,
  theta_bounds = c(bounds$theta_min, bounds$theta_max),
  lambda_bounds = c(bounds$lambda_min, bounds$lambda_max),
  alpha_bounds = c(bounds$alpha_min, bounds$alpha_max),
  delta_bounds = c(bounds$delta_min, bounds$delta_max),
  tau_bounds = c(bounds$tau_min, bounds$tau_max),
  k_bounds = c(bounds$k_min, bounds$k_max),
  omega_bounds = c(bounds$omega_min, bounds$omega_max),
  y0_bounds = c(bounds$y0_min, bounds$y0_max)
)
sample_model("flux_all_simpl_K", mdata_all)
```

### Fitted kurtosis - per site - no hyperdistribution

In this version, the kurtosis parameter $\kappa$ is fitted, with one value per site, and the productivity is therefore modelled as: 

$$ STPprod_{p,t} = \delta_p \cdot \left( \frac{t}{\tau_s} \cdot exp(1-\frac{t}{\tau_s}) \right)^{\kappa_s} $$

```{r}
#| message: false
#| warning: false
#| eval: false
ind_rec <- data_rec |>
  select(Site, Plot, sitenum, plotnum) |>
  unique() |>
  arrange(plotnum)

mdata_all <- list(
  n_equ = nrow(data_old),
  n_log = nrow(data_rec),
  s = max(data_old$sitenum),
  p_log = max(data_rec$plotnum),
  y_equ = data_old$stocks,
  in_equ = data_old$prod,
  out_equ = data_old$mort,
  y_log = data_rec$stocks,
  in_log = data_rec$prod,
  out_log = data_rec$mort,
  time = data_rec$trec,
  dt = data_rec$tdiff,
  site_equ = data_old$sitenum,
  site_log = data_rec$sitenum,
  plot_log = data_rec$plotnum,
  site_plot = ind_rec$sitenum,
  theta_bounds = c(bounds$theta_min, bounds$theta_max),
  lambda_bounds = c(bounds$lambda_min, bounds$lambda_max),
  alpha_bounds = c(bounds$alpha_min, bounds$alpha_max),
  delta_bounds = c(bounds$delta_min, bounds$delta_max),
  tau_bounds = c(bounds$tau_min, bounds$tau_max),
  k_bounds = c(bounds$k_min, bounds$k_max),
  omega_bounds = c(bounds$omega_min, bounds$omega_max),
  y0_bounds = c(bounds$y0_min, bounds$y0_max)
)
sample_model("flux_all_simpl_Ksite_nohyper", mdata_all)
```

### Fitted kurtosis - per site - with a hyperdistribution

In this version, the kurtosis parameter $\kappa$ is fitted, with one value per site, and the productivity is therefore modelled as: 

$$ STPprod_{p,t} = \delta_p \cdot \left( \frac{t}{\tau_s} \cdot exp(1-\frac{t}{\tau_s}) \right)^{\kappa_s} $$

We also added a hyperdistribution to all parameters $\kappa_s$: 

$$\kappa_s \sim \mathcal{N}(\mu_{\kappa}, \sigma_{\kappa})$$

```{r}
#| message: false
#| warning: false
#| eval: false
ind_rec <- data_rec |>
  select(Site, Plot, sitenum, plotnum) |>
  unique() |>
  arrange(plotnum)

mdata_all <- list(
  n_equ = nrow(data_old),
  n_log = nrow(data_rec),
  s = max(data_old$sitenum),
  p_log = max(data_rec$plotnum),
  y_equ = data_old$stocks,
  in_equ = data_old$prod,
  out_equ = data_old$mort,
  y_log = data_rec$stocks,
  in_log = data_rec$prod,
  out_log = data_rec$mort,
  time = data_rec$trec,
  dt = data_rec$tdiff,
  site_equ = data_old$sitenum,
  site_log = data_rec$sitenum,
  plot_log = data_rec$plotnum,
  site_plot = ind_rec$sitenum,
  theta_bounds = c(bounds$theta_min, bounds$theta_max),
  lambda_bounds = c(bounds$lambda_min, bounds$lambda_max),
  alpha_bounds = c(bounds$alpha_min, bounds$alpha_max),
  delta_bounds = c(bounds$delta_min, bounds$delta_max),
  tau_bounds = c(bounds$tau_min, bounds$tau_max),
  k_bounds = c(bounds$k_min, bounds$k_max),
  omega_bounds = c(bounds$omega_min, bounds$omega_max),
  y0_bounds = c(bounds$y0_min, bounds$y0_max)
)
sample_model("flux_all_simpl_Ksite", mdata_all)
```


## Results comparison

```{r}
#| message: false
#| warning: false
fit_flux_all <- c(
  "flux_all_simpl",
  "flux_all_simpl_K",
  "flux_all_simpl_Ksite_nohyper",
  "flux_all_simpl_Ksite"
) |>
  lapply(function(name) {
    as_cmdstan_fit(list.files(paste0("chains/", name),
      full.names = TRUE
    ))
  })
```

```{r}
#| message: false
#| warning: false
obs_data <- data_rec |>
  mutate(
    val = "obs",
    type = "rec"
  ) |>
  select(-plotnum) |>
  rbind(mutate(
    data_old,
    val = "obs",
    type = "equ"
  ))

pov_data <- lapply(1:4, function(i) {
  data_rec |>
    select(-plotnum) |>
    mutate(
      prod = fit_flux_all[[i]]$summary(c("mu_in_log"), median)$median,
      mort = fit_flux_all[[i]]$summary(c("mu_out_log"), median)$median,
      val = "pred",
      type = "rec"
    ) |>
    bind_rows(mutate(
      data_old,
      prod = fit_flux_all[[i]]$summary(c("theta_s"), median)$median[data_old$sitenum] * # nolint
        fit_flux_all[[i]]$summary(c("omega_s"), median)$median[data_old$sitenum], # nolint
      mort = fit_flux_all[[i]]$summary(c("theta_s"), median)$median[data_old$sitenum] * # nolint
        fit_flux_all[[i]]$summary(c("omega_s"), median)$median[data_old$sitenum], # nolint
      val = "pred",
      type = "equ"
    )) |>
    bind_rows(obs_data) |>
    pivot_longer(cols = c(prod, mort), names_to = "variable") |>
    pivot_wider(names_from = val, values_from = value) |>
    mutate(model = c("K=2", "One K", "Site K", "Site K + Hyperdistribution")[i])
}) |>
  bind_rows()
```

### Traceplots

```{r}
#| message: false
#| warning: false
lapply(fit_flux_all, function(x) {
  mcmc_trace(x$draws(variables = c("lp__"))) +
    theme(legend.position = "none")
}) %>% # nolint
  ggarrange(
    plotlist = .,
    labels = c("K=2", "One K", "Site K", "Site K + Hyperdistribution")
  )
```


### Goodness of fit

```{r}
#| message: false
#| warning: false
df_rmse <- pov_data |>
  group_by(variable, model) |>
  summarise(rmse = sqrt(mean((pred - obs)^2)))

labs <- c(prod = "AGB productivity", mort = "AGB mortality")

pov_data |>
  filter(variable == "prod") |>
  ggplot(aes(pred, obs)) +
  geom_point(alpha = 0.25, aes(col = Site, shape = type)) +
  geom_abline(col = "red", linetype = "dashed") +
  geom_text(
    data = filter(df_rmse, variable == "prod"),
    aes(
      x = -Inf, y = Inf, hjust = -0.1, vjust = 1,
      label = paste("RMSE =", round(rmse, 2))
    )
  ) +
  theme_bw() +
  xlab("Predicted") +
  ylab("Observed") +
  facet_wrap(~model, scales = "free", labeller = labeller(variable = labs))
```

### Predicted productivity

The predicted productivity trajectories are presented below.

```{r}
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 15
df_rmse <- pov_data |>
  filter(variable == "prod") |>
  group_by(model, Site) |>
  summarise(rmse = sqrt(mean((pred - obs)^2)))
ind_site <- data_rec |>
  select(Site, sitenum) |>
  unique() |>
  arrange(sitenum)
df_k <- lapply(1:4, function(i) {
  if (i == 1) {
    ind_site |>
      mutate(model = "K=2", k = "2")
  } else if (i == 2) {
    k <- round(fit_flux_all[[i]]$summary(c("k"))[, c("q5", "median", "q95")], 2)
    ind_site |>
      mutate(model = "One K", k = paste0(k$median, " [", k$q5, ",", k$q95, "]"))
  } else {
    fit_flux_all[[i]]$summary(c("k_s")) |>
      mutate(across(where(is.numeric), ~ round(.x, 2))) |>
      mutate(k = paste0(median, " [", q5, ",", q95, "]")) |>
      bind_cols(ind_site) |>
      mutate(model = c("Site K", "Site K + Hyperdistribution")[i - 2]) |>
      select(Site, sitenum, model, k)
  }
}) |>
  bind_rows()

pov_data |>
  subset(variable == "prod") |>
  mutate(Plot = plot_new) |>
  graph_traj(expression("AGB productivity [" ~ Mg ~ ha^-1 ~ yr^-1 ~ "]")) +
  geom_text(data = df_rmse, aes(
    x = -Inf, y = Inf, hjust = -0.1, vjust = 1,
    label = paste("RMSE =", round(rmse, 2))
  )) +
  geom_text(data = df_k, aes(
    x = -Inf, y = Inf, hjust = -0.1, vjust = 2, label = paste("K =", k)
  )) +
  facet_grid(Site ~ model, scales = "free")
```


## Conclusion

The model with one fitted kurtosis per site and no hyperdistribution predicted the productivity trajectories better. Our interpretation is that it introduces enough variability to capture the sites' differences in the productivity response, but the introduction of a hyperdistribution constrained the sites with little data too much, as the site with the most data (Paracou) had one of the lowest kurtosis values. 

We therefore chose this model. 
